---
title: "Take-home Exercise 2: Discovering impacts of COVID-19 on Thailand tourism economy at the province level using spatial and spatio-temporal statistics"
author: "Fu Wanqian"
published: "September 10, 2024"
modified: "September 22, 2024"
execute: 
  echo: true
  warning: false
  freeze: true
  cache: true
  eval: false
fotmat:
  html:
    code-fold: true
---

# 1 Set the Scene

Tourism is a major contributor to Thailand’s economy, making up about 20% of GDP. In 2019, the industry generated 90 billion US$, but due to the COVID-19 pandemic, this dropped to 24 billion US$ in 2020. Since September 2021, tourism revenues have been gradually recovering. However, the economic benefits are not evenly spread across the country, with key provinces like Bangkok, Phuket, Chiang Mai, Sukhothai, and Phetchaburi dominating the sector. In this exercise,we need to explore whether the tourism economy shows spatial and spatio-temporal dependence, and if so, identify clusters, outliers, and emerging hot or cold spots.

# 2 Setting the Analytical Tools

```{r}
pacman::p_load(sf, spdep, tmap, tidyverse, patchwork,spacetime)
```

# 3 Data Collection

Two data sets will be used in this exercise:

-   [Thailand Domestic Tourism Statistics](https://www.kaggle.com/datasets/thaweewatboy/thailand-domestic-tourism-statistics) at Kaggle. This dataset contains statistics on domestic tourism in Thailand from Jan 2019 to Feb 2023, broken down by province. The datasetincludes information on the number of tourists, the occupancy rate, and the profits generated by tourism in each province, as wel as thenationality of the tourists.

-   [Thailand - Subnational Administrative Boundaries](https://data.humdata.org/dataset/cod-ab-tha?) at HDX. This dataset has been used in the last exercise.

Use below codes to import these two raw datasets.

```{r}
thai_province <- st_read(dsn="data/geospatial", layer = 'tha_admbnda_adm1_rtsd_20220121')
```

```{r}
thai_tour <- read_csv("data/aspatial/thailand_domestic_tourism_2019_2023_ver2.csv")
```

And use below codes to generate a spatio-temporal dataset of thai tourism economy indicator.

```{r}
# change the column name to get identifier
colnames(thai_tour)[colnames(thai_tour) == "province_eng"] <- "ADM1_EN"
```

```{r}
# create year-month column
thai_tour$date <- as.Date(thai_tour$date)

thai_tour$year_month <- as.Date(format(thai_tour$date, "%Y-%m-01"))

```

```{r}
thai_tour$ADM1_EN <- tolower(thai_tour$ADM1_EN)
thai_province$ADM1_EN <- tolower(thai_province$ADM1_EN)
```

```{r}
thai_tour_spt <- spacetime(thai_tour, thai_province,
                       .loc_col = "ADM1_EN",
                       .time_col = "year_month")
```

```{r}
is_spacetime_cube(thai_tour_spt)
```

Since the spacetime object is sparse, we use below code to make it a full one.

```{r}
thai_tour_full <- complete_spacetime_cube(thai_tour_spt)
```

```{r}
dplyr::count(thai_tour_spt, year_month, ADM1_EN)
```

```{r}
is_spacetime_cube(thai_tour_spt)
```
